---
title: "Data Visualization using ggplot2"
date: "January 2023"
abstract: |
  This is adapted from content in the book *R for Data Science*.

  https://r4ds.had.co.nz/data-visualisation.html
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Getting started with ggplot

The "gg" in ggplot2 stands for "grammar of graphics", a coherent system for describing and building graphs.  We will use the mpg dataframe to explore ggplot capabilities.  Type ?mpg or help(mpg) to get info on the dataset.

```{r prereq}
library(ggplot2)
library(dplyr)
```

Note that displ is the car's engine size and hwy is the mpg for the car on the highway. ggplot call creates axes on which to plot the selected data.  we then add a layer of geometric points that represent the x and y coordinates that we want to show. 

```{r firstplot}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

In-class Exercises:

1. Run ggplot(data = mpg)  What do you see? Why?
2. What does the drv variable describe?  Read the help for ?mpg to find out.
3. How many rows are in mtcars?  How many columns?
4. Make a scatterplot of hwy versus cyl.

## Aesthetic mappings

An aesthetic is a visual property of the objects in your plot.  Includes shape, size, color of the points.  We can set the aesthetic (aes) values to depend on a variable value. If a property does not convey information about a variable, but just changes its appearance, it is set outside aes() but inside the mapping.

```{r aes}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy,  color = class),size = 3)
```
Experiment with setting different shapes (choose a number from 0-20).  Size is given in mm.  Color is a character string of a color.


Exercises:

1) Which variables in mpg are categorical? Which are continuous? You can type ?mpg or view(mpg) for help.

2) Map a continuous variable to color, size, and shape.  How do these aesthetics behave differently for categorical versus continuous variables?

3) Can you map the same variables to multiple aesthetics?


\newpage

### Common Problem
Make sure the + is at the end of the line.

## Facets


Split data into separate plots - one for each category of a categorical variable.  Use facet_wrap() which needs a formula for an argument.  Formulas (equations) are given using ~ ; variable passed to facet_wrap should be discrete

```{r facets}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y= hwy)) +
  facet_wrap(~class, nrow=2)
```
\newpage

Use facet_grid to plot on a combination of two variables

```{r facetgrid}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y= hwy)) +
  facet_grid(drv~cyl)
```
Exercises:

1) What happens if you facet on a continuous variable?

2) What do empty cells in the previous plot signal?

3) What does . do?  For example, try these:

      ggplot(data=mpg) + geom_point(mapping=aes(displ,hwy)) + facet_grid(drv ~ .)
      
      ggplot(data=mpg) + geom_point(mapping=aes(displ,hwy)) + facet_grid(. ~ cyl)
      
4) Read ?facet_wrap.  What does nrow do?  What does ncol do? Why doesn't facet_grid have these options?    
      
\newpage
## Geometric Objects

A geom is a geometrical object that a plot uses to represent the data.  Examples include bars, lines, points, boxplots, etc.
Let's try changing our geom in the above examples from point to smooth.
```{r smooth}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
\newpage
Note that not every aesthetic works with every geom.  For instance, shape works for points, but not lines and linetype works for lines and not points.
```{r smoothlinetype}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
\newpage
Maybe this looks strange without showing the points.  See if you can add the points into the graph and color code both the points and the lines according to drv.
```{r smoothpoints}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color=drv)) +
  geom_point(mapping = aes(x=displ, y=hwy, color = drv))
```

\newpage
However, this was repetitious.  We could instead set aesthetics for all the geoms globally in the ggplot function and then use any aesthetics in specific geoms to overwrite or add to what we specified globally.  Try it.
```{r smoothpoint2}
ggplot(data = mpg, mapping =aes(x=displ, y=hwy, color = drv) ) +
  geom_smooth(mapping = aes(linetype = drv)) +
  geom_point()
```
\newpage
We could add just one trend line to the data if we want.  Here is how to add just the trend line for front wheel drive cars
```{r smoothpoint3}
ggplot(data = mpg, mapping =aes(x=displ, y=hwy, color = drv) ) +
  geom_point() +
  geom_smooth(data = filter(mpg, drv == "f"), se = FALSE)
```
\newpage
Exercises:

1) What geom would you use for a line chart? A boxplot? A histogram? 

2) Run this code and discuss:
```{r}
  ggplot(data=mpg, mapping=aes(displ, hwy, color=drv)) + geom_point() +     
      geom_smooth(se=FALSE)
```  
\newpage
3) Are the following two bits of code equivalent?
```{r}
  ggplot() +
    geom_point(data=mpg, mapping = aes(displ, hwy)) +
    geom_smooth(data=mpg, mapping = aes(displ, hwy))
```
\newpage
```{r}
  ggplot(data=mpg, mapping=aes(displ,hwy)) + geom_point() + geom_smooth()
```  
\newpage  

## Statistical Transformations

Some geoms use multiple rows to create an intermediate result which is then plotted.  For instance, bar chart uses that statistical function count and then plots the results.  Similarly, histograms and frequency polygons also bin your data and then count the number of rows that fall into each bin.  Smoothers fit a model to the data and then plot predictions. Boxplots use a statistical summary of the data and then plot the results. 

Stat (statistical transformation) is the way the new (intermediate) values are calculated.  Look at ?geom_bar and look for the section on stat.  Each stat has a default geom and each geom has a default stat, so they can generally be interchanged if desired.

```{r stat}
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut))
```

\newpage
```{r stat2}
ggplot(data=diamonds) +
  stat_count(mapping=aes(x=cut))
```
\newpage
If you have the actual values that you want to plot already in a table, you can use the identity stat to pick up those values.
```{r fakedata}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
demo
ggplot(data=demo)+
  geom_bar(mapping=aes(x=cut, y=freq), stat = "identity")
```

\newpage
If instead you want to see the proportion, try overriding the stat value.
```{r statprop}
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut, y = ..prop.., group = 1))
```
\newpage
Check out the ggplot2 cheatsheet for a list of stats that you can use.  To see what a particular one does (e.g. stat_bin), use ? (like ?stat_bin)  Look up stat_summary and see what it does.  How does that map to the following code?
```{r statsum}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
\newpage

Exercises for in class: (3.7.1 in book)

1) What is the default geom associated wit stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

2) What does geom_co() do? How is it different from geom_bar()?

3) Most geoms and stats come in pairs that are almost always used together. Read the docs and make a list of the pairs (or find a list). What do they have in common?

4) What variable does stat_smooth() compute? What parameters control its behavior?

5) In our proportion bar chart, we need to set group=1. Why? What is the problem with these graphs below?

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, y=after_stat(prop)))
```

\newpage

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut, fill=color, y=after_stat(prop)))
```


\newpage 

## Position Adjustments

We can change the colors of our barplots.  If you use the color in the mapping, it will just change the line color.  To change the color of the whole bar, use fill instead.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))
```

\newpage

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x=cut, fill = cut))
```
\newpage
If you use a different variable for the fill aesthetic, you get stacked bars.  This is because the default for position adjustment is stacking. 

```{r stack}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x=cut, fill = clarity))
```
\newpage
The other 3 options for position are Identity, fill and dodge.  Let's see what each one does.
```{r}
ggplot(data = diamonds, 
       mapping = aes(x=cut, color = clarity)) +
  geom_bar(fill=NA, position = "identity")
```
\newpage
```{r}
ggplot(data = diamonds, 
       mapping = aes(x=cut, fill = clarity)) +
  geom_bar(position = "fill")
```
\newpage
```{r}
ggplot(data = diamonds, 
       mapping = aes(x=cut, fill = clarity)) +
  geom_bar(position = "dodge")
```
\newpage
One final position adjustment is sometimes useful for points, and that is jitter.

```{r jitter}
ggplot(data = mpg, 
       mapping = aes(x = displ, y = hwy)) +
  geom_point()
```

\newpage

```{r}
ggplot(data = mpg, 
       mapping = aes(x = displ, y = hwy)) +
  geom_point(position = "jitter")
```
\newpage

## Coordinate Systems

Default is the Cartesian coordinate system with x and y positions acting independently to determine the location of each point.

A few other coordinate systems include 
* coord_flip (flips x and y axes)
* coord_quickmap (sets aspect ratio for maps)
* coord_polar (uses polar coordinates)

```{r}
boxes <- ggplot(data = mpg, 
                mapping=aes(x=class, y =hwy))
boxes + geom_boxplot()
```
\newpage

```{r}
boxes + geom_boxplot() + coord_flip()
```
\newpage
```{r}
bar <- ggplot(data = diamonds ) +
  geom_bar(
    mapping = aes(x=cut, fill = cut),
    show.legend=FALSE,
    width = 1) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar + coord_flip()
```
\newpage
```{r}

bar + coord_polar()
```

Note: Just because you can do a thing, don't assume it is a good idea.  What issues do you see with this graphic!?